---
title: "Inferential analysis of ToothGrowth dataset"
author: "Piotr Turski"
date: "September 23, 2015"
output: pdf_document
---

## Overview

This document shows analysis of ToothGrowth dataset version `r packageVersion('datasets')` provided by R.

## Exploratory data analyses


```{r}
data(ToothGrowth)
str(ToothGrowth)
```

According to `?ToothGrowth`, dataset content:
"The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid)". Newer version's (3.3.0) description says that ascorbic acid is coded as VC. 

A plot suggested in the documentation helps to understand the data:
```{r , fig.height = 4}
coplot(len ~ dose | supp, data = ToothGrowth, panel = panel.smooth,
       xlab = "ToothGrowth data: length vs dose, given type of supplement")
```

## Compare tooth growth by supp and dose

Let's do `t.test` to compare: supplement types, dosage, supplement type per each dose 

### Compare supplement type
```{r}
t.test(len ~ supp, data = ToothGrowth)
```

Confidence interval contains 0 so we cannot reject null hypothesis that supplement type doesn't matter.

### Compare dosage

```{r message=F}
library(dplyr)
t.test(len ~ dose, data = filter(ToothGrowth, dose != 2))
```

Confidence interval doesn't contain 0 so we reject null hypothesis that dose 0.5mg vs 1mg doesn't matter. Mean length for dose 1mg is greater.

```{r}
t.test(len ~ dose, data = filter(ToothGrowth, dose != 0.5))
```

Similary, dose 2mg gives larger length mean than dose 1mg.

### Compare supplement type per dosage

```{r}
t.test(len ~ supp, data = filter(ToothGrowth, dose == 0.5))
```

Confidence interval doesn't contain 0 so we reject null hypothesis that supplement type for dose 0.5mg doesn't matter. Mean length for dose 1mg is greater for orange juice.

```{r}
t.test(len ~ supp, data = filter(ToothGrowth, dose == 1))
```

Confidence interval doesn't contain 0 so we reject null hypothesis that supplement type for dose 1mg doesn't matter. Mean length for dose 1mg is greater for orange juice.

```{r}
t.test(len ~ supp, data = filter(ToothGrowth, dose == 2))
```

Confidence interval contains 0 so we cannot reject null hypothesis that supplement type for dose 2mg doesn't matter.

## Conclusions

* Tooth length increases with dose.
* For doses 0,5mg and 1mg orange juice causes greater tooth length than ascorbic acid.
* For dose 2mg and irrespective of the dose, supplement type doesn't matter.

## Assumptions

* Distribution of means is approximately normal
* Samples are random and independent
* Tested animals are representative for the whole population